Mathematical appendix to "Spatio-temporal estimates of HIV risk group proportions for adolescent

### 2022-06-18 src/docs_paper/appendix.Rmd
Improve MP transformation section, move constraints section, correct some indexing issues (not using K anymore), improve survey weighting section (still needs more though) #128

We index the four risk groups by $k \in \{1, 2, 3, 4\}$ and denote being in either the third or fourth risk group by $k = 3^{+}$.
Let $c \in \{1, \ldots, 13\}$ denote the 13 studied AGYW priority countries, each partitioned into districts according to Table A.
For each district $i \in \{1, \ldots, n\}$, year $t \in \{1999, \ldots, 2018\}$, age category $a \in \{\text{15--19}, \text{20--24}, \text{25--29}\}$, and risk group $k \in \{1, 2, 3^{+}\}$ we consider the multinomial random vector
\begin{equation}
    \y_{ita} = (y_{ita1}, \ldots, y_{ita3^{+}})^\top \sim \text{Multinomial}(m_{ita}; \, p_{ita1}, \ldots, p_{ita3^{+}}),
\end{equation}
where the number of women in risk group $k$ is $y_{itak}$, the sample size is $m_{ita} = \sum_{k = 1}^{3^{+}} y_{itak}$, and $p_{itak} > 0$ is the probability of membership of the $k$th risk group with $\sum_{k = 1}^{3^{+}} p_{itak} = 1$.

To efficiently model this data we used the multinomial-Poisson transformation, which allows multinomial logistic regression models to be cast as equivalent Poisson log-linear models [@baker1994multinomial] of the form
\begin{align}
    y_{itak} &\sim \text{Poisson}(\kappa_{itak}), \label{eq:poisson} \\
    \log(\kappa_{itak}) &= \eta_{itak}, \label{eq:linearpredictor}
\end{align}
for suitable choice of the linear predictor $\eta_{itak}$, which we describe in more detail in Section.
Then, to estimate the proportion of those in the $k = 3^{+}$ risk group that were respectively in the $k = 3$ and $k = 4$ risk groups, we fit a separate logistic regression model of the form
\begin{align}
    y_{ia4} &\sim \text{Binomial} \left( y_{ia3} + y_{ia4}, q_{ia} \right), \\
    q_{ia} &= p_{ia4} / (p_{ia3} + p_{ia4}) = p_{ia4} / p_{ia{3^+}}, \\
    q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), 
\end{align}
again for suitable choice of the linear predictor $\eta_{ia}$, which we describe in more detail in Section.
This approach allowed us to include all surveys in the multinomial regression model, but only those surveys with a specific transactional sex question to estimate the proportions $\{p_{ia4}\}$.
As all such surveys occurred in 2013-2018, in the logistic regresion model $q_{ia}$ was assumed to be constant over time.

To allow uncertainty quantification, we took 1000 posterior samples, indexed by $s$, from each of the multinomial $\{p_{itak}^s\}$ and logistic regression $\{q_{ia}^s\}$ models.
Samples $p_{ita3}^s$ and $p_{ita4}^s$ were then generated by splitting samples from the $k = 3^{+}$ category of the multinomial model via $p_{ita3}^s = (1 - q_{ia}^s)p_{ita{3^+}^s}$ and $p_{ita4}^s = q_{ia}^s p_{ita{3^+}}^s$.

@@

The transformation reframes a given multinomial logisitic regression model as an equivalent Poisson log-linear model of the form
\begin{align}
    y_{itak} &\sim \text{Poisson}(\kappa_{itak}), \\
    \log(\kappa_{itak}) &= \eta_{itak}.
\end{align}
Category probabilities are obtained by applying the softmax function to the linear predictor as follows
\begin{equation}
    p_{itak} = \frac{\exp(\eta_{itak})}{\sum_{k = 1}^{3^{+}} \exp(\eta_{itak})} = \frac{\kappa_{itak}}{\sum_{k = 1}^{3^{+}} \kappa_{itak}}.
\end{equation}
In the equivalent model, the sample sizes $m_{ita} = \sum_k y_{itak}$ are treated as random, rather than fixed as they would be in the multinomial logistic regression model, and follow a Poisson distribution
\begin{equation}
    m_{ita} \sim \text{Poisson} \left( \kappa_{ita} \right),
\end{equation}
where the intensity $\kappa_{ita} = \sum_k \kappa_{itak}$ is given by a sum of intensities over each category.
The basis of the Poisson trick is that, conditional on their sum, Poisson counts are jointly multinomially distributed [@mccullagh1989generalized]
\begin{equation}
    \mathbf{y}_{ita} \, | \, m_{ita} \sim \text{Multinomial} \left( m_{ita}; \frac{\kappa_{ita1}}{\kappa_{ita}}, \ldots, \frac{\kappa_{itaK}}{\kappa_{ita}} \right).
\end{equation}
As such, the joint distribution of $p(\mathbf{y}_{ita}, m_{ita})$ is
\begin{align}
p(m_{ita}) p(\mathbf{y}_{ita} \, | \, m_{ita}) 
&= \exp(-\kappa_{ita}) \frac{(\kappa_{ita})^{m_{ita}}}{m_{ita}!} \times \frac{m_{ita}!}{\prod_k y_{itak}!} \prod_k \left( \frac{\kappa_{itak}}{\kappa_{ita}} \right)^{y_{itak}} \\
&= \prod_k \left( \frac{\exp(-\kappa_{itak}) \left( \kappa_{itak} \right)^{y_{itak}}}{y_{itak}!} \right) \\
&= \prod_k \text{Poisson} \left( y_{itak} \, | \, \kappa_{itak} \right). \label{eq:prodpoisson}
\end{align}
Equation corresponds to the product of $K$ independent Poisson likelihoods, and the joint distribution over all the data $\{\mathbf{y}_{ita}, m_{ita}\}_{i, t, a}$ is given by $\prod_{i, t, a, k} \text{Poisson} \left( y_{itak} \, | \, \kappa_{itak} \right)$.
The Poisson log-linear model, which vicariously includes the sample sizes as random, is equivalent only when these normalisation constants are recovered exactly by the model.
To ensure that this is the case, we included observation-specific random effects $\theta_{ita}$ in the equation for the linear predictor in every model we considered.
Multiplying each of $\{\kappa_{itak}\}_{k = 1}^K$ by $\exp(\theta_{ita})$ has no effect on the multinomial probabilities, but does provide the flexibility required for $\kappa_{ita}$ to exactly recover $m_{ita}$ as desired.
Although in theory an improper prior $\theta_{ita} \propto 1$ should be used, in practise, by keeping $\eta_{ita}$ small so that arbitrarily large values of the observation-specific random effects $\theta_{ita}$ are not required, it is sufficient (and practically preferable for inference) to instead use a vague prior.
As such, we set $\theta_{ita} \sim \mathcal{N}(0, \tau_\theta^{-1})$, with fixed precision $\tau_\theta = 0.000001$ (or a standard deviation of $\sigma_\theta = 1000$).

@@

We considered eight models (Table \ref{tab:multinomial-models}) for $\eta_{ita}$ of the form
$$
\eta_{ita} = \theta_{ita} + \beta_k + \alpha_{ak} + \zeta_{ck} + \phi_{ik} + \gamma_{tk} + \delta_{itk}.
$$
Observation random effects $\theta_{ita}$ with a diffuse prior, $\mathcal{N}(0, 1000^2)$, were included in every model to allow the sample sizes $m_{ita}$ to be recovered exactly, as required for equivalency of the models.
To capture country-specific proportion estimates for each category $k = 1, 2, 3^{+}$, we included category random effects $\beta_k \sim \mathcal{N}(0, \tau_\beta^{-1})$ and country-category random effects $\zeta_{ck} \sim \mathcal{N}(0, \tau_\zeta^{-1})$.
We allowed heterogeneity in risk group proportions by age by including age-country-category random effects $\alpha_{ack} \sim \mathcal{N}(0, \tau_\alpha^{-1})$.
We considered two specifications, independent and identically distributed (IID) and Besag, for the space-category $\phi_{ik}$ random effects and two specifications, IID and first order autoregressive (AR1), for the year-category $\gamma_{tk}$ random effects, as well as separable space-year-category interactions $\delta_{itk}$ with corresponding Kronecker product structure.
All random effect precision parameters $\tau \in \{\tau_\beta, \tau_\alpha, \tau_\zeta, \tau_\phi, \tau_\gamma\, \tau_\delta\}$ were given independent penalised complexity (PC) priors [@simpson2017penalising] with base model $\sigma = 0$ given by $p(\tau) = 0.5 \nu \tau^{-3/2} \exp \left( - \nu \tau^{-1/2} \right)$
where $\nu = - \ln(0.01) / 2.5$ such that  $\mathbb{P}(\sigma > 2.5) = 0.01$.

To ensure interpretable posterior inferences of random effect contribution, we applied sum-to-zero constraints such that the country-category, age-country-category, space-category and year-category random effects did not alter overall category probabilities.
For the space-year-category random effects, we analogously applied sum-to-zero constraints to maintain roles of the space-category and year-category random effects.
Together, these were\footnote{The constraints for the space-year-category random effects were unable to be implemented fully, though this only effects the variance components interpretation.}:
\begin{enumerate}
    \item Category $\sum_k \beta_k = 0$
    \item Country $\sum_c \zeta_{ck} = 0, \, \forall \, k$
    \item Age-country $\sum_a \alpha_{ack} = 0, \, \forall \, k \quad \text{TODO}$,
    \item Spatial $\sum_i \phi_{ik} = 0, \, \forall \, k$
    \item Temporal $\sum_t \gamma_{tk} = 0, \, \forall \, k$
    \item Spatio-temporal $\sum_i \delta_{itk} = 0, \, \forall \, t, k;
\quad \sum_t \delta_{itk} = 0, \, \forall \, i, k;
\quad \sum_k \delta_{itk} = 0, \, \forall \, i, t$
\end{enumerate}

@@

The specifications we considered for the space-category random effects were IID $\phi_{ik} \sim \mathcal{N}(0, \tau_\phi^{-1})$ and Besag grouped by category
$$
\bphi = (\phi_{11}, \ldots, \phi_{n1}, \ldots, \phi_{1K}, \ldots \phi_{nK})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\phi \mathbf{R}^\star_\phi)^{-})
$$
where the scaled structure matrix $\mathbf{R}^\star_\phi = \mathbf{R}^\star_b \otimes \mathbf{I}$ is given by the Kronecker product of the scaled Besag structure matrix $\mathbf{R}^\star_b$ and the identity matrix $\mathbf{I}$.
Scaling of the structure matrix to have generalised variance one ensures interpretable priors may be placed on the precision parameter [@sorbye2014scaling].
We follow the further recommendations of @freni2018note with regard to disconnected adjacency graphs, singletons and constraints.
The Besag structure matrix $\mathbf{R}_b$ is obtained by the precision matrix of the random effects $\mathbf{b} = (b_1, \ldots, b_n)^\top$ with the full conditionals
\begin{equation}
b_i \, | \, \mathbf{b}_{-i} \sim \N\left(\frac{\sum_{j: j \sim i} b_j}{n_{\delta i}}, \frac{1}{n_{\delta i}}\right),
\end{equation}
where $j \sim i$ if the districts $A_i$ and $A_j$ are adjacent, and $n_{\delta i}$ is the number of districts adjacent to $A_i$.

@@

The specifications we considered for the survey-category random effects were IID $\phi_{tk} \sim \mathcal{N}(0, \tau_\phi^{-1})$ and AR1 grouped by category
$$
\bm{\gamma} = (\gamma_{11}, \ldots, \gamma_{1K}, \ldots, \gamma_{T1}, \ldots, \gamma_{TK})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\phi \mathbf{R}^\star_\gamma)^{-})
$$
where the scaled structure matrix $\mathbf{R}^\star_\gamma = \mathbf{R}^\star_r \otimes \mathbf{I}$ is given by the Kronecker product of a scaled AR1 structure matrix $\mathbf{R}^\star_r$ and the identity matrix $\mathbf{I}$.
The AR1 structure matrix $\mathbf{R}_r$ is obtained by precision matrix of the random effects $\mathbf{r} = (r_1, \ldots, r_T)^\top$ specified by
\begin{align}
r_1 &\sim \left( 0, \frac{1}{1 - \rho^2} \right), \\
r_t &= \rho r_{t - 1} + \epsilon_t, \quad t = 2, \ldots, T, 
\end{align}
where $\epsilon_t \sim \mathcal{N}(0, 1)$ and $|\rho| < 1$.
For the lag-one correlation parameter $\rho$, we used the PC prior, as derived by @sorbye2017penalised, with base model $\rho = 1$ and condition $\mathbb{P}(\rho > 0 = 0.75)$.
We chose a base model corresponding to no change in behaviour over time, rather than the alternative $\rho = 0$ corresponding to no correlation in behaviour over time, as we judged the former to be more plausible a priori.

@@

For each model including spatio-temporal random effects we used the specification
\begin{equation}
    \bm{\delta} = (\delta_{111}, \ldots, \delta_{nTK})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\delta \mathbf{R}^\star_\delta)^{-}),
\end{equation}
where $\mathbf{R}^\star_\delta$ is a Kronecker product of the relevant space, time and category structure matrices.
In particular, for each of the four models including spatio-temporal interactions, these matrices are
\begin{itemize}
\item
1x: No spatial or temporal structure (Type I) $\mathbf{R}^\star_\delta = \mathbf{I} \otimes \mathbf{I} \otimes \mathbf{I}$,
\item
2x: Structured spatial and unstructured temporal (Type II) $\mathbf{R}^\star_\delta = \mathbf{R}^\star_b \otimes \mathbf{I} \otimes \mathbf{I}$,
\item
3x: Unstructured spatial and structured temporal (Type III) $\mathbf{R}^\star_\delta = \mathbf{I} \otimes \mathbf{R}^\star_a \otimes \mathbf{I}$,
\item
4x: Structured spatial and temporal (Type IV) $\mathbf{R}^\star_\delta = \mathbf{R}^\star_b \otimes \mathbf{R}^\star_a \otimes \mathbf{I}$,
\\
\end{itemize}
where the first, second and third elements of the product represent space, time and category respectively, and the interaction type provided in brackets is given according to the @knorr2000bayesian framework.

@@

The surveys we used utilise a complex design, in which each individual has an unequal probability of being included in the sample.
For example, the DHS often employs a two-stage cluster design, first taking an urban rural stratified sample of ennumeration areas, before selecting households from each ennumeration area using systematic sampling.
To account for the survey design, we use a weighted pseudo-likelihood where the observed counts $y_{itak}$ are replaced by effective counts $y^\star_{itak}$ calculated using survey weights $w_{itak}$.
We multiplied direct estimates produced using the `survey` package [@JSSv009i08] by the Kish effective sample size [@kish1965survey]
\begin{equation}
    m^\star_{itak} = \left(\sum_{itak} w_{itak} \right)^2 / \sum_{itak} {w_{itak}}^2
\end{equation}
to obtain $y^\star_{itak}$.
As these counts may not be integers, the Poisson likelihood in Equation~\eqref{eq:poisson} is inappropriate.
Instead, we used a generalised Poisson pseudo-likelihood $y^\star \sim \text{xPoisson}(\kappa)$, as implemented by `family = "xPoisson"` in `R-INLA`, given by
\begin{equation}
    p(y^\star) = \frac{\kappa^{y^\star}}{\left \lfloor{y^\star!}\right \rfloor } \exp \left(- \kappa \right).
\end{equation}

@@

We considered six logistic regression models (Table~\ref{tab:logistic-models}) each including a constant $\beta_0 \sim \mathcal{N}(-2, 1^2)$, age random effects $\alpha_a \sim \mathcal{N}(0, \tau_\alpha^{-1})$ and country random effects $\zeta_c \sim \mathcal{N}(0, \tau_\zeta^{-1})$.
We considered two specifications (IID, Besag) for the spatial random effects $\phi_i$.
To aid estimation with sparse data, we also considered national-level covariates for the proportion of men who have paid for sex ever \texttt{cfswever} or in the last twelve months \texttt{cfswrecent}, available from @hodgins2022population.
For both random effect precision parameters $\tau \in \{\tau_\alpha, \tau_\zeta\}$ we used the PC prior with base model $\sigma = 0$ and $\mathbb{P}(\sigma > 2.5 = 0.01)$.
For the regression parameters $\beta \in \{\beta_\texttt{cfswever}, \beta_\texttt{cfswrecent}\}$ we used the prior $\beta \sim \mathcal{N}(0, 2.5^2)$.

>>>>

Let $c \in \{1, \ldots, 13\}$ denote the 13 studied AGYW priority countries, partitioned into subnational units $i \in \{1, \ldots, n\}$ according to Table B.5.
We considered the years 1999-2018 denoted as $t \in \{1, \ldots, T\}$ and age groups $a \in \{\text{15--19}, \text{20--24}, \text{25--29}\}$
Let the four risk groups be $k \in \{1, 2, 3, 4\}$ and denote being in either the third or fourth risk group by $k = 3^{+}$.
We modelled the probability of being in each risk group $k \in \{1, 2, 3^{+}\}$ using a multinomial logisitic regression model.
Typically, this model would be specified using a multinomial likelihood
\begin{equation}
    \y_{ita} = (y_{ita1}, \ldots, y_{ita3^{+}})^\top \sim \text{Multinomial}(m_{ita}; \, p_{ita1}, \ldots, p_{ita3^{+}}),
\end{equation}
where the number of women in risk group $k$ is $y_{itak}$, the sample size is $m_{ita} = \sum_{k = 1}^{3^{+}} y_{itak}$, and $p_{itak} > 0$ is the probability of membership of the $k$th risk group with $\sum_{k = 1}^{3^{+}} p_{itak} = 1$.
Following the baseline category approach with $k = 1$ say, linear predictors would then be specified for the additive log ratios $\log(p_{itak} / p_{ita1})$.
Instead of taking this approach, to facilitate inference in the `R-INLA`, we used an equivalent Poisson log-linear model via the multinomial-Poisson transformation [@baker1994multinomial].
This transformation, and further details about this model, are presented in Section 1.2.

Next, we fit a logistic regression model to estimate the proportion of those in the $k = 3^{+}$ risk group that were in the $k = 3$ and $k = 4$ risk groups respectively.
This model was of the form
\begin{align}
    y_{ia4} &\sim \text{Binomial} \left( y_{ia3} + y_{ia4}, q_{ia} \right), \\
    q_{ia} &= p_{ia4} / (p_{ia3} + p_{ia4}) = p_{ia4} / p_{ia{3^+}}, \\
    q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), 
\end{align}
for suitable choice of the linear predictor $\eta_{ia}$, described in Section 1.3.
Taking this modular approach allowed us to include all surveys in the multinomial regression model, but only those surveys with a specific transactional sex question to estimate the proportions $\{p_{ia4}\}$.
As all such surveys occurred in 2013-2018, in the logistic regresion model $q_{ia}$ was assumed to be constant over time.

To facilitate uncertainty quantification, we took 1000 posterior samples, indexed by $s$, from each of the multinomial $\{p_{itak}^s\}$ and logistic regression $\{q_{ia}^s\}$ models.
Samples $p_{ita3}^s$ and $p_{ita4}^s$ were then generated by splitting samples from the $k = 3^{+}$ category of the multinomial model via $p_{ita3}^s = (1 - q_{ia}^s)p_{ita{3^+}^s}$ and $p_{ita4}^s = q_{ia}^s p_{ita{3^+}}^s$.

@@

The multinomial-Poisson transformation reframes a given multinomial logisitic regression model as an equivalent Poisson log-linear model of the form
\begin{align}
    y_{itak} &\sim \text{Poisson}(\kappa_{itak}), \label{eq:poisson} \\
    \log(\kappa_{itak}) &= \eta_{itak},
\end{align}
for certain choice of the linear predictor $\eta_{itak}$.
The basis of the transformation is that, conditional on their sum, Poisson counts are jointly multinomially distributed [@mccullagh1989generalized] as follows
\begin{equation}
    \mathbf{y}_{ita} \, | \, m_{ita} \sim \text{Multinomial} \left( m_{ita}; \frac{\kappa_{ita1}}{\kappa_{ita}}, \ldots, \frac{\kappa_{ita3^{+}}}{\kappa_{ita}} \right),
\end{equation}
where $\kappa_{ita} = \sum_{k = 1}^{3^{+}} \kappa_{itak}$ such that category probabilites are obtained by the softmax funciton
\begin{equation}
    p_{itak} = \frac{\exp(\eta_{itak})}{\sum_{k = 1}^{3^{+}} \exp(\eta_{itak})} = \frac{\kappa_{itak}}{\sum_{k = 1}^{3^{+}} \kappa_{itak}} = \frac{\kappa_{itak}}{\kappa_{ita}}.
\end{equation}
In the equivalent model, the sample sizes $m_{ita} = \sum_k y_{itak}$ are treated as random, rather than fixed as they would be in the multinomial logistic regression model, taking a Poisson distribution
\begin{equation}
    m_{ita} \sim \text{Poisson} \left( \kappa_{ita} \right).
\end{equation}
In the equivalent model, the joint distribution of $p(\mathbf{y}_{ita}, m_{ita}) = p(\mathbf{y}_{ita} \, | \, m_{ita})p(m_{ita})$ is
\begin{align}
p(\mathbf{y}_{ita}, m_{ita}) &= \exp(-\kappa_{ita}) \frac{(\kappa_{ita})^{m_{ita}}}{m_{ita}!} \times \frac{m_{ita}!}{\prod_k y_{itak}!} \prod_k \left( \frac{\kappa_{itak}}{\kappa_{ita}} \right)^{y_{itak}} \\
&= \prod_k \left( \frac{\exp(-\kappa_{itak}) \left( \kappa_{itak} \right)^{y_{itak}}}{y_{itak}!} \right) \\
&= \prod_k \text{Poisson} \left( y_{itak} \, | \, \kappa_{itak} \right). \label{eq:prodpoisson}
\end{align}
corresponding to the product of independent Poisson likelihoods as in Equation \ref{eq:poisson}.
This model, including random sample sizes, is equivalent to the multinomial logisitc regression only when these normalisation constants are recovered exactly.
To ensure that this is the case, one approach is to include observation-specific random effects $\theta_{ita}$ in the equation for the linear predictor.
Multiplying each of $\{\kappa_{itak}\}_{k = 1}^{3^+}$ by $\exp(\theta_{ita})$ has no effect on the category probabilities, but does provide the necessary flexibility for $\kappa_{ita}$ to recover $m_{ita}$ exactly.
Although in theory an improper prior $\theta_{ita} \propto 1$ should be used, in practise, by keeping $\eta_{ita}$ otherwise small using appropriate constraints, so that arbitrarily large values of $\theta_{ita}$ are not required, it is sufficient (and practically preferable for inference) to instead use a vague prior.

@@

We considered eight models (Table \ref{tab:multinomial-models}) for $\eta_{ita}$ of the form
$$
\eta_{ita} = \theta_{ita} + \beta_k + \alpha_{ak} + \zeta_{ck} + \phi_{ik} + \gamma_{tk} + \delta_{itk}.
$$
As described above, we included observation random effects in all models we considered, and set $\theta_{ita} \sim \mathcal{N}(0, \tau_\theta^{-1})$, with fixed precision $\tau_\theta = 0.000001$ (or a standard deviation of $\sigma_\theta = 1000$).
To capture country-specific proportion estimates for each category, we included category random effects $\beta_k \sim \mathcal{N}(0, \tau_\beta^{-1})$ and country-category random effects $\zeta_{ck} \sim \mathcal{N}(0, \tau_\zeta^{-1})$.
Heterogeneity in risk group proportions by age was allowed by including age-country-category random effects $\alpha_{ack} \sim \mathcal{N}(0, \tau_\alpha^{-1})$.
We considered two specifications, independent and identically distributed (IID) and Besag, for the space-category $\phi_{ik}$ random effects (Section \ref{sec:spatial-random-effects}) and two specifications, IID and first order autoregressive (AR1), for the year-category $\gamma_{tk}$ random effects (Section \ref{sec:temporal-random-effects}).
We also considered separable space-year-category interactions $\delta_{itk}$ with corresponding Kronecker product structure (Section \ref{sec:spatio-temporal-random-effects}).

All random effect precision parameters $\tau \in \{\tau_\beta, \tau_\alpha, \tau_\zeta, \tau_\phi, \tau_\gamma\, \tau_\delta\}$ were given independent penalised complexity (PC) priors [@simpson2017penalising] with base model $\sigma = 0$ given by $p(\tau) = 0.5 \nu \tau^{-3/2} \exp \left( - \nu \tau^{-1/2} \right)$ where $\nu = - \ln(0.01) / 2.5$ such that  $\mathbb{P}(\sigma > 2.5) = 0.01$.

@@

The specifications we considered were IID $\phi_{ik} \sim \mathcal{N}(0, \tau_\phi^{-1})$ and Besag grouped by category
$$
\bphi = (\phi_{11}, \ldots, \phi_{n1}, \ldots, \phi_{13^{+}}, \ldots \phi_{n3^{+}})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\phi \mathbf{R}^\star_\phi)^{-})
$$
where the scaled structure matrix $\mathbf{R}^\star_\phi = \mathbf{R}^\star_b \otimes \mathbf{I}$ is given by the Kronecker product of the scaled Besag structure matrix $\mathbf{R}^\star_b$ and the identity matrix $\mathbf{I}$.
Scaling of the structure matrix to have generalised variance one ensures interpretable priors may be placed on the precision parameter [@sorbye2014scaling].
We follow the further recommendations of @freni2018note with regard to disconnected adjacency graphs, singletons and constraints.
The Besag structure matrix $\mathbf{R}_b$ is obtained by the precision matrix of the random effects $\mathbf{b} = (b_1, \ldots, b_n)^\top$ with the full conditionals
\begin{equation}
b_i \, | \, \mathbf{b}_{-i} \sim \N\left(\frac{\sum_{j: j \sim i} b_j}{n_{\delta i}}, \frac{1}{n_{\delta i}}\right),
\end{equation}
where $j \sim i$ if the districts $A_i$ and $A_j$ are adjacent, and $n_{\delta i}$ is the number of districts adjacent to $A_i$.

@@

The specifications we considered were IID $\phi_{tk} \sim \mathcal{N}(0, \tau_\phi^{-1})$ and AR1 grouped by category
$$
\bm{\gamma} = (\gamma_{11}, \ldots, \gamma_{13^{+}}, \ldots, \gamma_{T1}, \ldots, \gamma_{T3^{+}})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\phi \mathbf{R}^\star_\gamma)^{-})
$$
where the scaled structure matrix $\mathbf{R}^\star_\gamma = \mathbf{R}^\star_r \otimes \mathbf{I}$ is given by the Kronecker product of a scaled AR1 structure matrix $\mathbf{R}^\star_r$ and the identity matrix $\mathbf{I}$.
The AR1 structure matrix $\mathbf{R}_r$ is obtained by precision matrix of the random effects $\mathbf{r} = (r_1, \ldots, r_T)^\top$ specified by
\begin{align}
r_1 &\sim \left( 0, \frac{1}{1 - \rho^2} \right), \\
r_t &= \rho r_{t - 1} + \epsilon_t, \quad t = 2, \ldots, T, 
\end{align}
where $\epsilon_t \sim \mathcal{N}(0, 1)$ and $|\rho| < 1$.
For the lag-one correlation parameter $\rho$, we used the PC prior, as derived by @sorbye2017penalised, with base model $\rho = 1$ and condition $\mathbb{P}(\rho > 0 = 0.75)$.
We chose a base model corresponding to no change in behaviour over time, rather than the alternative $\rho = 0$ corresponding to no correlation in behaviour over time, as we judged the former to be more plausible a priori.

@@

We used the specification
\begin{equation}
    \bm{\delta} = (\delta_{111}, \ldots, \delta_{nT3^{+}})^\top \sim \mathcal{N}(\mathbf{0}, (\tau_\delta \mathbf{R}^\star_\delta)^{-}),
\end{equation}
where $\mathbf{R}^\star_\delta$ is a Kronecker product of the relevant space, time and category structure matrices.
In particular, for each of the four models including spatio-temporal interactions, these matrices are
\begin{itemize}
\item
1x: No spatial or temporal structure (Type I) $\mathbf{R}^\star_\delta = \mathbf{I} \otimes \mathbf{I} \otimes \mathbf{I}$,
\item
2x: Structured spatial and unstructured temporal (Type II) $\mathbf{R}^\star_\delta = \mathbf{R}^\star_b \otimes \mathbf{I} \otimes \mathbf{I}$,
\item
3x: Unstructured spatial and structured temporal (Type III) $\mathbf{R}^\star_\delta = \mathbf{I} \otimes \mathbf{R}^\star_a \otimes \mathbf{I}$,
\item
4x: Structured spatial and temporal (Type IV) $\mathbf{R}^\star_\delta = \mathbf{R}^\star_b \otimes \mathbf{R}^\star_a \otimes \mathbf{I}$,
\\
\end{itemize}
where the first, second and third elements of the product represent space, time and category respectively, and the interaction type provided in brackets is given according to the @knorr2000bayesian framework.

#### Constraints

To ensure interpretable posterior inferences of random effect contribution, we applied sum-to-zero constraints such that the country-category, age-country-category, space-category and year-category random effects did not alter overall category probabilities.
For the space-year-category random effects, we analogously applied sum-to-zero constraints to maintain roles of the space-category and year-category random effects.
Together, these were:
\begin{enumerate}
    \item Category $\sum_k \beta_k = 0$
    \item Country $\sum_c \zeta_{ck} = 0, \, \forall \, k$
    \item Age-country $\sum_a \alpha_{ack} = 0, \, \forall \, k \quad \text{TODO}$,
    \item Spatial $\sum_i \phi_{ik} = 0, \, \forall \, k$
    \item Temporal $\sum_t \gamma_{tk} = 0, \, \forall \, k$
    \item Spatio-temporal\footnote{The constraints for the space-year-category random effects were unable to be implemented fully, though this only effects the variance components interpretation.} $\sum_i \delta_{itk} = 0, \, \forall \, t, k;
\quad \sum_t \delta_{itk} = 0, \, \forall \, i, k;
\quad \sum_k \delta_{itk} = 0, \, \forall \, i, t$
\end{enumerate}

@@

The surveys we included use a complex design, in which each individual has an unequal probability of being included in the sample.
For example, the DHS often employs a two-stage cluster design, first taking an urban rural stratified sample of ennumeration areas, before selecting households from each ennumeration area using systematic sampling [@measure2012sampling].
To account for the survey design, we use a weighted pseudo-likelihood where the observed counts $y$ are replaced by effective counts $y^\star$ calculated using the survey weights $w_j$ of all individuals $j$ in the corresponding strata.
We multiplied direct estimates produced using the `survey` package [@JSSv009i08] by the Kish effective sample size [@kish1965survey]
\begin{equation}
    m^\star = \frac{\left(\sum_j w_j \right)^2}{\sum_j {w_j}^2}
\end{equation}
to obtain $y^\star$.
These counts may not be integers, and the Poisson likelihood we used in Equation \ref{eq:poisson} is inappropriate.
Instead, we used a generalised Poisson pseudo-likelihood $y^\star \sim \text{xPoisson}(\kappa)$, given by
\begin{equation}
    p(y^\star) = \frac{\kappa^{y^\star}}{\left \lfloor{y^\star!}\right \rfloor } \exp \left(- \kappa \right),
\end{equation}
as implemented by `family = "xPoisson"` in `R-INLA`.

To fill.

@@

We considered six logistic regression models (Table \ref{tab:logistic-models}) each including a constant $\beta_0 \sim \mathcal{N}(-2, 1^2)$, age random effects $\alpha_a \sim \mathcal{N}(0, \tau_\alpha^{-1})$ and country random effects $\zeta_c \sim \mathcal{N}(0, \tau_\zeta^{-1})$.
We considered two specifications (IID, Besag) for the spatial random effects $\phi_i$.
To aid estimation with sparse data, we also considered national-level covariates for the proportion of men who have paid for sex ever \texttt{cfswever} or in the last twelve months \texttt{cfswrecent}, available from @hodgins2022population.
For both random effect precision parameters $\tau \in \{\tau_\alpha, \tau_\zeta\}$ we used the PC prior with base model $\sigma = 0$ and $\mathbb{P}(\sigma > 2.5 = 0.01)$.
For the regression parameters $\beta \in \{\beta_\texttt{cfswever}, \beta_\texttt{cfswrecent}\}$ we used the prior $\beta \sim \mathcal{N}(0, 2.5^2)$.
